Distributed Range-Only Localisation that Preserves Sensor and Navigator Privacies

### 2022-10-11 authors_responses/conditional_accept_2/authors_response.tex
Added remaining changes from reviewer comments Updated authors response with information from updated TAC response (different editor now)

Dear Dr Zhiwei Gao,\\
Dear Reviewers,

@@

\section*{Response to the Editor's Report}
\def\thesection{E}
\begin{rebuttal} %\setcounter{enumi}{-1}
\item \reviewtext{I am pleased to inform you that the paper is acceptable for publication in the Transactions provided that you can make the modifications described below.}

We are glad to hear the positive response, thank you. Below we respond to the remaining reviewer comments.

@@

Thank you for noticing this. The error has now been corrected.

@@

We regret that the intended meaning was not clearer. The two expressions were used interchangeably due to sensors being the only corruptible parties by an attacker. The section has been updated to explicitly state that attackers are honest-but-curious and that sensors can be corrupted. We hope this clarifies the intended meaning.

@@

TODO respond.

@@

TODO respond.

>>>>

Dear Prof. David Castanon,\\
Dear Reviewers,

@@

\section*{Response to the Editor's Report}
\def\thesection{E}
\begin{rebuttal} %\setcounter{enumi}{-1}
\item \reviewtext{I am pleased to inform you that the paper is acceptable for publication in the Transactions provided that you can make the modifications described below.

The reviewers would like the authors to double check the symbols and formulas and clarify some concerns on sensor attacks etc, which are detailed in their comments.}

We are glad to hear the positive response, thank you. Below we clarify the remaining concerns from reviewer comments.

@@

The error has now been corrected, thank you for noticing it.

@@

This was not written clearly and we are grateful that it was pointed out. The section has been updated to explicitly state that attackers are honest-but-curious and that the navigator or sensors can be corrupted. We hope this clarifies the intended meaning.

@@

The reviewer is correct to notice the inconsistency with notations in this section. Key subscripts have been added to make it clear that the encryptions $\mathcal{E}(\cdot)$ are performed with the keys generated in the $\mathsf{Setup}(\cdot)$ algorithm and the linear combinations $l_i^{(t)}$ modified to denote the unencrypted sum.

@@

Sorry about the mistake, the missing year has been added and the remaining references inspected for additional errors.

### 2022-10-11 main.tex
Added remaining changes from reviewer comments Updated authors response with information from updated TAC response (different editor now)

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted by $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|e}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including timestep $e$. When discussing cryptography, $\mathcal{E}_{sk}(\cdot)$ and $\mathcal{D}_{sk}(\cdot)$ will denote encryption and decryption with key $sk$, respectively, with $sk$ ommited when inferrable from context. $\mathbb{Z}_N$ is used for the set of integers modulo $N$ and $\mathbb{Z}^*_N$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ the binary concatination operator, and the term \textit{negligible function} refers to its cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

At every \textit{instance} $t$ (used to distinguish from an estimation \textit{timestep}), the navigator first broadcasts $m$ weights $\omega_j^{(t)}, j\in\{1,\dots,m\}$ to all sensors $i\in\{1,\dots,n\}$, who individually compute linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{j,i}^{(t)}\omega_i^{(t)}$ based on their measurement data $a_{j,i}$. Linear combinations are then sent back to the navigator, who computes their sum $\sum^n_{i=1}l^{(t)}_{i}$. This two-step linear combination aggregation protocol has been visually displayed in figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize,scale=0.95]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and Combination Step};
    % Navigator
    \fill (3.25,4.75) [pyplotblue!70] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.9225) {plane};
    % Sensors
    \node at (1.5,1.375) {Sensor $1$};
    \fill [pyplotorange!70] (0.25,1.625) rectangle (2.875,2.625);
    \node at (1.625,2.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,1.375) {Sensor $n$};
    \fill [pyplotorange!70] (3.625,1.625) rectangle (6.25,2.625);
    \node at (5,2.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,2.125) circle (0.05);
    \fill [black] (3,2.125) circle (0.05);
    \fill [black] (3.25,2.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4.25) (5,2.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4.25) (1.5,2.875)};
    \fill [lightgray] (2.125,3.875) rectangle (4.375,3.125);
    \node at (3.25,3.5) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

Before giving a formal definition for the construction and security of our desired encryption scheme, we make the following assumptions on the capabilities of the participants.
\begin{LaTeXdescription}
    \item[Global Navigator Broadcast] We assume that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.
    \item[Consistent Navigator Broadcast] We assume that broadcast information from the navigator is received equally by all sensors. This means the navigator may not send different weights to individual sensors during a single instance $t$.
    \item[Honest-but-Curious Sensors] We adopt the honest-but-curious attacker model for all involved sensors, meaning that they follow the localisation procedure correctly but may store or use any gained sensitive information.
\end{LaTeXdescription}
We justify the global broadcast assumption by noting that any subset of sensors within the range of the navigator can be considered a group and treated as the global set during estimation, generalising the method, while the wide-spread use of cheap non-directional antennas supports the assumption of consistent broadcasts. The final assumption refers to the known problem of misbehaving sensors \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}, often requiring additional complicated detection mechanisms, and will not be considered in this work.

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator, and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}(\omega_1^{(t)}),\dots,\mathcal{E}(\omega_m^{(t)}), a^{(t)}_{i,1},\dots,a^{(t)}_{i,m})$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination denoted $l^{(t)}_i = \mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ using its secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, l^{(t)}_1,\dots,l^{(t)}_n)$] At instance $t$, the navigator computes the aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix~\ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learnt by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\section{Conclusion} \label{sec:conclusion}
We have presented a localisation filter in the presence of range-only sensors, which preserves both navigator and sensor privacies. A suitable cryptographic scheme has been introduced and a filter implementation compared and evaluated. Privacy-preserving range-only localisation is suitable for use in environments where sensor networks are untrusted or location is considered private and we hope to extend the method to broader measurement models in the future. Additional future work includes exploring more computationally efficient encryption schemes, the security implications of sensors that are not only honest-but-curious and expanding the LCAO notion to enforce the consistent broadcast assumption.

>>>>

\subsection{Notation}
Throughout this work, we make use of the following notation. Underlined characters, $\vec{a}$, denote vectors, uppercase bold characters, $\mat{A}$, denote matrices. $\mat{A}^{-1}$ is the matrix inverse while $\mat{A}^\top$ is its transpose. The expected value of a random variable is denoted by $\mathsf{E}[\cdot]$, while the variance of a random scalar and covariance of a random vector by $\mathsf{Var}[\cdot]$ and $\mathsf{Cov}[\cdot]$, respectively. $|a|$ denotes absolute value, $\lVert\vec{a}\rVert$ the vector norm, $\{\dots\}$ is used for sets and $\langle\dots\rangle$ for ordered sequences. When estimating a state, notation $\hat{\vec{x}}_{k|e}$ will denote an estimate of $\vec{x}$ at timestep $k$ given measurements from timesteps up to and including timestep $e$. When discussing cryptography, $\mathcal{E}_{sk}(\cdot)$ and $\mathcal{D}_{sk}(\cdot)$ will denote encryption and decryption with key $sk$, respectively, with $sk$ ommited when inferrable from context. $\mathbb{Z}_N$ is used for the set of integers modulo $N$ and $\mathbb{Z}^*_N$ for its multiplicative group. $\mathsf{lcm}(\cdot,\cdot)$ is the lowest common denominator, $\|$ is the binary concatination operator, and the term \textit{negligible function} refers to its cryptographic definition in \cite{katzIntroductionModernCryptography2008}.

@@

At every \textit{instance} $t$ (used to distinguish from an estimation \textit{timestep}), the navigator first broadcasts $m$ weights $\omega_j^{(t)}, j\in\{1,\dots,m\}$ to all sensors $i\in\{1,\dots,n\}$, who individually compute linear combinations $l^{(t)}_i=\sum^m_{j=1}a_{i,j}^{(t)}\omega_j^{(t)}$ based on their measurement data $a_{j,i}$. Linear combinations are then sent back to the navigator, who computes their sum $\sum^n_{i=1}l^{(t)}_{i}$. This two-step linear combination aggregation protocol has been visually displayed in figure \ref{fig:agg_steps}.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[font=\footnotesize,scale=0.95]
    % Step 1
    \node at (3.25,5.5) {1. Broadcast and Combination Step};
    % Navigator
    \fill (3.25,4.75) [pyplotblue!70] ellipse (0.4 and 0.4);
    \pic[xscale=0.22,yscale=0.3] at (3.25,4.9225) {plane};
    % Sensors
    \node at (1.5,1.375) {Sensor $1$};
    \fill [pyplotorange!70] (0.25,1.625) rectangle (2.875,2.625);
    \node at (1.625,2.125) {$\displaystyle l_1^{(t)} \!=\! \sum^m_{j=1}a_{1,j}^{(t)}\omega_j^{(t)}$};
    \node at (5,1.375) {Sensor $n$};
    \fill [pyplotorange!70] (3.625,1.625) rectangle (6.25,2.625);
    \node at (5,2.125) {$\displaystyle l_n^{(t)} \!=\! \sum^m_{j=1}a_{n,j}^{(t)}\omega_j^{(t)}$};
    \fill [black] (3.5,2.125) circle (0.05);
    \fill [black] (3,2.125) circle (0.05);
    \fill [black] (3.25,2.125) circle (0.05);
    % Lines
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3.5,4.25) (5,2.875)};
    \draw [-latex] plot[smooth, tension=.7] coordinates {(3,4.25) (1.5,2.875)};
    \fill [lightgray] (2.125,3.875) rectangle (4.375,3.125);
    \node at (3.25,3.5) {$\langle\omega_1^{(t)},\dots ,\omega_m^{(t)}\rangle$};

@@

Before giving a formal definition for the construction and security of our desired encryption scheme, we make the following assumptions on the capabilities of the participants.
\begin{LaTeXdescription}
    \item[Global Navigator Broadcast] We assume that broadcast information from the navigator is received by \textit{all} sensors involved in the protocol.
    \item[Consistent Navigator Broadcast] We assume that broadcast information from the navigator is received equally by all sensors. This means the navigator may not send different weights to individual sensors during a single instance $t$.
    \item[Honest-but-Curious Attackers] We adopt the honest-but-curious attacker model for all corrupted parties that an attacker controls, meaning that the navigator and sensors follow the localisation procedure correctly but attackers may store or use any gained sensitive information.
\end{LaTeXdescription}
We justify the global broadcast assumption by noting that any subset of sensors within the range of the navigator can be considered a group and treated as the global set during estimation, generalising the method, while the wide-spread use of cheap non-directional antennas supports the assumption of consistent broadcasts. The final assumption refers to the known problem of misbehaving sensors \cite{lazosSeRLocSecureRangeindependent2004,ben-galOutlierDetection2005}, often requiring additional complicated detection mechanisms, and will not be considered in this work.

We are now ready to define the type of encryption scheme we want for the specified communication protocol and the security guarantees it should provide. We let a linear combination aggregation scheme be defined as a tuple of the four algorithms $(\mathsf{Setup}, \mathsf{Enc}, \mathsf{CombEnc}, \mathsf{AggDec})$. These will be used by a trusted setup party, the navigator, and sensors $i\in\{1,\dots,n\}$. They are defined as follows.
\begin{LaTeXdescription}
    \item[$\mathsf{Setup}(\kappa)$] On input of security parameter $\kappa$, generate public parameters $\mathsf{pub}$, the number of weights $m$, the navigator's public and private keys $pk_0$ and $sk_0$ and the sensor private keys $sk_i,\,i\in\{1,\dots,n\}$.
    \item[$\mathsf{Enc}(pk_0, x)$] The navigator and sensors can encrypt any value $x$ with the navigator's public key $pk_0$ and obtain the encryption $\mathcal{E}_{pk_0}(x)$.
    \item[$\mathsf{CombEnc}(t, pk_0, sk_i, \mathcal{E}_{pk_0}(\omega_1^{(t)}),\!...,\mathcal{E}_{pk_0}(\omega_m^{(t)}),a^{(t)}_{i,1},\!...,a^{(t)}_{i,m})\hspace*{1.2pt}$] At instance $t$, sensor $i$ computes and obtains the encrypted linear combination $\mathcal{E}_{pk_0,sk_i}(\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j)$ with secret key $sk_i$.
    \item[$\mathsf{AggDec}(t, pk_0, sk_0, \mathcal{E}_{pk_0,sk_1}(l^{(t)}_1),\dots,\mathcal{E}_{pk_0,sk_n}(l^{(t)}_n))$] At instance $t$, the navigator computes the plaintext aggregation of linear combinations $\sum^{n}_{i=1}l_i^{(t)}$ using its public and private keys $pk_0$, $sk_0$.
\end{LaTeXdescription}
The security notions we want these algorithms to meet reflect the previously stated estimation privacy goals. The navigator should learn no information from individual sensors while sensors should learn no information from the navigator or any other sensors. In the context of the introduced communication protocol, this can be summarised as the following notions.
\begin{LaTeXdescription}
    \item[Indistinguishable Weights] No colluding subset of sensors gains any new knowledge about the navigator weights $\omega^{(t)}_j,\,j\in\{1,\dots,m\}$ when receiving only their encryptions from the current and previous instances and having the ability to encrypt plaintexts of their choice.
    \item[Linear Combination Aggregator Obliviousness] No colluding subset \textit{excluding} the navigator gains additional information about the remaining sensor values to be weighted, $a^{(t)}_{i,j},\,j\in\{1,\dots,m\}$, where sensor $i$ is not colluding, given only encryptions of their linear combinations $l_i$ from the current and previous instances. Any colluding subset \textit{including} the navigator learns only the sum of all linear combinations weighted by weights of their choice, $\sum^{n}_{i=1}l_i^{(t)}=\sum^{n}_{i=1}\sum^{m}_{j=1} a^{(t)}_{i,j}\omega^{(t)}_j$.
\end{LaTeXdescription}
While indistinguishable weights can be achieved by encrypting weights with an encryption scheme meeting the notion of Indistinguishability under the Chosen Plaintext Attack (IND-CPA) \cite{katzIntroductionModernCryptography2008}, the novel notion of Linear Combination Aggregator Obliviousness (LCAO) has been formalised as a typical cryptographic game between attacker and challenger in appendix~\ref{app:lcao}. Lastly, we conclude the cryptographic problem definition with the following important remark.
\begin{remark}
    A leakage function including weights from the navigator requires extra care to be taken when giving its definition. If an attacker compromises the navigator, they have control over the weights, and therefore the leakage function. We note that in the leakage function above, $\sum^n_{i=1}\sum^m_{j=1}a^{(t)}_{i,j}\omega^{(t)}_j$, an individual sum weighted by the same weight may be learnt by an attacker, \textit{e.g.}, $\sum^n_{i=1}a^{(t)}_{i,1}$ given weights $(1,0,\dots,0)$, but that individual sensor values $a^{(t)}_{i,j}$ remain private due to the assumption of a consistent broadcast.
\end{remark}

@@

\section{Conclusion} \label{sec:conclusion}
We have presented a localisation filter in the presence of range-only sensors, which preserves both navigator and sensor privacies. A suitable cryptographic scheme has been introduced and a filter implementation compared and evaluated. Privacy-preserving range-only localisation is suitable for use in environments where sensor networks are untrusted or location is considered private and we hope to extend the method to broader measurement models in the future. Additional future work includes exploring more computationally efficient encryption schemes, the security implications of attackers that are not only honest-but-curious and expanding the LCAO notion to enforce the consistent broadcast assumption.
